Foregrounds in Wide--Field Redshifted 21~cm Power Spectra

### 2015-02-09 ms.tex
Change title and Introduction slightly
from 3e35322:
============

In ms.tex,

(*) Change title and last paragraph of Introduction section
(*) Change wording in Pitchfork subsection slightly
(*) Change wording slightly in caption to figure on foreground
mitigation by baseline screening

\shorttitle{EoR Foregrounds in Fourier Space}
\shortauthors{Thyagarajan et~al.}

@@

%% \title{Detecting Epoch of Reionization in redshifted 21~cm: contamination in EoR window}
%% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
\title{Foregrounds in Fourier Delay Space for Wide--Field Redshifted 21~cm Observatories}

@@

{\bf We investigate signatures generic to all wide--field measurements of EoR power spectra and explore antenna shapes that are most desirable from the viewpoint of foreground contamination for future EoR experiments}. After matching observations from the Murchison Widefield Array with instrumental simulations of the entire sky, we examine the {\bf shape of the delay spectrum} of each baseline. We report two important findings: in the {\it avoidance} strategy, the foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. We quantify these by separating the simulations into components based on type of foreground emission. We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

In \S\ref{sec:delay-spectrum} we {\bf provide a brief overview of the delay spectrum technique} \citep{par12a,par12b}. We establish the signatures arising from the nature of wide--field measurements in \S\ref{sec:wide-field}. In \S\ref{sec:sim}, we present the foreground model and a variety of instrument models to rank antenna shapes based on the criterion of foreground contamination. In \S\ref{sec:MWA}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are compared with the data and analyzed for foreground signatures. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

The peak at $\tau=0$ (corresponding to the primary lobe in the power pattern) with a value of $\sim 10^7$--$10^8$~K$^2$(Mpc/$h$)$^3$, independent of antenna spacing, is predominantly determined by compact emission. The corresponding peak at zero--delay from diffuse foreground model is $\sim 10^3$ times fainter and decreases rapidly with increase in antenna spacing. This is the response expected from different antenna spacings towards compact and diffuse emission.

Near the horizon delay limits, the diffuse component is brighter relative to the compact component. Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at zero--delay. In fact, even on widely spaced antennas, diffuse emission in the {\bf delay power spectrum} near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude. This feature is described in \S\ref{sec:diffuse-long-baselines}. We attribute this to wide--field measurement effects discussed in \S\ref{sec:wide-field}.

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
% \includegraphics[width=\linewidth]{f13.eps}
% \includegraphics[width=\linewidth]{fig14.eps}
\includegraphics[width=\linewidth]{figure11.eps}
\caption{Drop in foreground contamination in the {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines. Eastward baselines with varying degrees of directedness -- $|\theta_\textrm{b}|<$~7.5\arcdeg (solid circles), $|\theta_\textrm{b}|<$~15\arcdeg~ (solid squares), and $|\theta_\textrm{b}|<$~22.5\arcdeg (solid stars) -- and lengths $|\boldsymbol{b}| \le |\boldsymbol{b}|_\textrm{max}$ ($x$--axis) are discarded. Loss of data (dotted lines) is measured by discarded baselines as a fraction of the total number for the corresponding cases. Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$2 ($|\theta_\textrm{b}|\le$~7.5\arcdeg) to $\sim$100 ($|\theta_\textrm{b}|\le$~22.5\arcdeg). The latter limit can be achieved with a mere 5\% loss of data at $|\boldsymbol{b}|_\textrm{max}\simeq30$~m, and discarding longer baselines ($|\boldsymbol{b}|\gtrsim 30$~m) has no effect in further reducing foreground contamination. \label{fig:screening}}
\end{figure}

>>>>

\shorttitle{Foregrounds in EoR Power Spectra}
\shortauthors{Thyagarajan et~al.}

@@

%% \title{Detecting Epoch of Reionization in redshifted 21~cm: contamination in EoR window}
%% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
\title{Foregrounds in Wide--Field Redshifted 21~cm Power Spectra}

@@

% {\bf We investigate signatures generic to all wide--field measurements of EoR power spectra and explore antenna shapes that are most desirable from the viewpoint of foreground contamination for future EoR experiments}. After matching observations from the Murchison Widefield Array with instrumental simulations of the entire sky, we examine the {\bf shape of the delay spectrum} of each baseline. We report two important findings: in the {\it avoidance} strategy, the foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. We quantify these by separating the simulations into components based on type of foreground emission. We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

{\bf In \S\ref{sec:delay-spectrum} we provide a brief overview of the delay spectrum technique \citep{par12a,par12b}. We investigate signatures generic to all wide--field measurements of EoR power spectra in \S\ref{sec:wide-field}. In \S\ref{sec:sim}, we present the foreground model and a variety of instrument models to rank antenna shapes based on foreground contamination. In \S\ref{sec:MWA}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are compared with the data and analyzed for foreground signatures. We report two important findings: foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique that minimizes foreground contamination using prior knowledge of the sky to down--weight adversely contaminated baselines}. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

The peak at $\tau=0$ (corresponding to the primary lobe in the power pattern) with a value of $\sim 10^7$--$10^8$~K$^2$(Mpc/$h$)$^3$, independent of antenna spacing, is predominantly determined by compact emission. The peak at $\tau=0$ from diffuse foreground model is $\sim 10^3$ times fainter and decreases rapidly with increase in antenna spacing. This is the response expected from different antenna spacings towards compact and diffuse emission.

Near the horizon delay limits, the diffuse component is brighter relative to the compact component. Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at $\tau=0$. In fact, even on widely spaced antennas, diffuse emission in the {\bf delay power spectrum} near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude. This feature was described in \S\ref{sec:diffuse-long-baselines}. We attribute this to wide--field measurement effects discussed in \S\ref{sec:wide-field}.

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
% \includegraphics[width=\linewidth]{f13.eps}
% \includegraphics[width=\linewidth]{fig14.eps}
\includegraphics[width=\linewidth]{figure11.eps}
\caption{Drop in foreground contamination in the MWA {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines. Eastward baselines with varying degrees of directedness -- $|\theta_\textrm{b}|<$~7.5\arcdeg (solid circles), $|\theta_\textrm{b}|<$~15\arcdeg~ (solid squares), and $|\theta_\textrm{b}|<$~22.5\arcdeg (solid stars) -- and lengths $|\boldsymbol{b}| \le |\boldsymbol{b}|_\textrm{max}$ ($x$--axis) are discarded. Loss of data (dotted lines) is measured by discarded baselines as a fraction of the total number for the corresponding cases. Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$2 ($|\theta_\textrm{b}|\le$~7.5\arcdeg) to $\sim$100 ($|\theta_\textrm{b}|\le$~22.5\arcdeg). The latter limit can be achieved with a mere 5\% loss of data at $|\boldsymbol{b}|_\textrm{max}\simeq30$~m, and discarding longer baselines ($|\boldsymbol{b}|\gtrsim 30$~m) has no effect in further reducing foreground contamination. \label{fig:screening}}
\end{figure}
